\documentclass[a4paper,11pt]{exam}

\usepackage[T1]{fontenc}
\usepackage[top=2cm, bottom=2cm, left=3cm, right=3cm]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{bm} % bold greek
\usepackage{tcolorbox}
\usepackage{xcolor}
\usepackage{forloop}
\usepackage{amsbsy}
%\usepackage[framed,numbered]{matlab-prettifier}
%\usepackage{filecontents}
\usepackage[numbers,sort&compress,comma]{natbib}

\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg,.mps,.eps,.ps}
\graphicspath{{figs/}}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\newcommand*{\clapp}[1]{\hbox to 0pt{\hss#1\hss}}
\newcommand*{\mat}[1]{\boldsymbol{\mathrm{#1}}}
\newcommand*{\subdims}[3]{\clapp{\raisebox{#1}[0pt][0pt]{$\scriptstyle(#2 \times #3)$}}}
\fboxrule=1pt

\newcommand{\defvec}[1]{\expandafter\newcommand\csname v#1\endcsname{{\mathbf{#1}}}}
\newcommand{\defrv}[1]{\expandafter\newcommand\csname rv#1\endcsname{{\mathbf{#1}}}}
\newcounter{ct}
\forLoop{1}{26}{ct}{
    \edef\letter{\alph{ct}}
    \expandafter\defvec\letter
}
% distinguish random vector
\forLoop{1}{26}{ct}{
    \edef\letter{\Alph{ct}}
    \expandafter\defrv\letter
}
\forLoop{1}{26}{ct}{
    \edef\letter{\alph{ct}}
    \expandafter\defrv\letter
}
\newcommand{\fundamentalSolution}{\boldsymbol{\Phi}}
\newcommand{\tvv}[1]{\mathbf{\tilde{#1}}} % tilde vectors
\newcommand{\hvv}[1]{\mathbf{\hat{#1}}} % hat vectors

\newcommand{\braket}[2]{\ensuremath{\langle#1\mid#2\rangle}}
\newcommand{\bra}[1]{\ensuremath{\langle#1|}}
\newcommand{\ket}[1]{\ensuremath{|#1\rangle}}
\newcommand{\horzbar}{\rule[.5ex]{1.5em}{0.4pt}}

\newcommand{\inv}{^{-1}}
\newcommand{\trp}{{^\top}} % transpose
\newcommand{\ctrp}{{^\dagger}} % transpose
\newcommand{\tp}[1]{\ensuremath{{#1}^\top}} % transpose
\newcommand{\dm}[1]{\ensuremath{\mathrm{d}{#1}}} % dx dy dz dmu
\newcommand{\RN}[2]{\frac{\dm{#1}}{\dm{#2}}} % (Radon-Nikodym) derivative
\newcommand{\PD}[2]{\frac{\partial #1}{\partial #2}} % partial derivative
\newcommand{\norm}[1]{\ensuremath{\Vert{#1}\Vert}}
\providecommand{\abs}[1]{\lvert#1\rvert}
\newcommand{\field}[1]{\ensuremath{\mathbb{#1}}}
\newcommand{\Kfield}{\field{K}}
\newcommand{\reals}{\field{R}}
\newcommand{\complex}{\field{C}}
\newcommand{\integers}{\field{Z}}
\newcommand{\naturalNumbers}{\field{N}}

\DeclareMathOperator*{\E}{\mathbb{E}} % expectation
\DeclareMathOperator*{\var}{var}
\DeclareMathOperator*{\cov}{cov}
\DeclareMathOperator*{\diag}{diag}
\newcommand{\system}[2]{\mathcal{#1}\left[ #2 \right]}
\DeclareMathOperator*{\tr}{tr} % trace
\newcommand{\indicator}[1]{\mathbf{1}_{#1}} % indicator function
\newcommand{\onehalf}{\frac{1}{2}}

\newcommand{\iid}{{\it i.i.d.}\xspace}
\newcommand{\iidsample}{\stackrel{iid}{\sim}}
\newcommand{\normalDist}{\mathcal{N}}
\newcounter{homework}
\newcommand{\homework}{\stepcounter{homework}\textcolor{violet}{\textbf{Homework \thehomework:}~}}
\newcommand{\funfact}{\textbf{Fun Fact:}~}
\newcommand{\myparagraph}[1]{\paragraph{#1}\mbox{}\\}


\pagestyle{headandfoot}
\runningheadrule
\runningheader{INDP 2023}{Statistics and Data Analysis for Biological Sciences}{Il Memming Park}

\title{Day 1 - The goat is more likely to be behind the other door}
\author{Il Memming Park, Hyungju Jeon}

\begin{document}
\maketitle
\begin{tcolorbox}[colback=black!1!,title=Learning Objective 1]
   Apply basic rules of probability. Determine when random variables are independent or dependent. Interpret distribution plots.
\end{tcolorbox}

\subsection{Conditional Probability}

\myparagraph{The problem of false positives} No diagnostic test is perfect in the real world. 
For example, in a flow cytometry test, we may have issues with unspecific binding and specific cross-reactivity with off targets resulting in false positives, as well as lack of reactivity with the stated target antigen resulting in false negatives. Mathematically, these can be described using \textbf{sensitivity} and \textbf{specificity}. 
\begin{itemize}
    \item Sensitivity is the probability of a positive test result, conditioned on the individual truly being positive.
    \item Specificity is the probability of a negative test result, conditioned on the individual truly being negative.
\end{itemize}

Suppose we have a test designed to identify the presence of a pathogen in patient blood samples. The sensitivity of the test is 80\% and the specificity is 99\%.  

\begin{questions}
    \question Suppose that 10\% of the population is infected with the pathogen.
    \begin{parts}
        \part What is the probability that a randomly selected patient will be tested positive?
        \vspace{\stretch{0.3}}
        \part If a randomly selected patient is tested and the test result is \textbf{positive}, what is the probability that the patient, in fact, has the pathogen?
        \vspace{\stretch{0.3}}
        \part If a randomly selected patient is tested and the test result is \textbf{negative}, what is the probability that the patient, in fact, has the pathogen?
        \vspace{\stretch{0.3}}
    \end{parts}
    \question Suppose now that 0.1\% of the population is infected with the pathogen. Does the probabilty above change? If so, how?
\end{questions}

\myparagraph{Monty Hall puzzle} The Monty Hall problem is a (in)famous probability puzzle, loosely based on the American television game show \textit{Let's Make a Deal} and named after its original host, \textit{Monty Hall}.

Suppose you're on a game show, and you're given the choice of three doors:
\begin{itemize}
    \item Behind one door is a car; behind the others, goats.
    \item The player pick a door. Then, the host Monty Hall, who knows what's behind the doors, will open one of the doors that has not been chosen, thus revealing a goat (he obviously won't open the car door).
    \item He then says to you, "Do you want to switch to the other door?"
\end{itemize}
\begin{questions}
    \question{What is the probability of winning the car if you stick with original choice?}
    \vspace{\stretch{0.3}}
    \question{What is the probability of winning the car if you switch?}
    \vspace{\stretch{0.3}}
\end{questions}


\myparagraph{The Two Children puzzle} Original problem is as follows
\begin{itemize}
    \item Mr. Jones has two children.  
    \item What is the probability that both children are girls?
    \item Now, if at least one of them is a boy. What is the probability that both children are boys?
\end{itemize}

\homework Simulate Monty Hall puzzle and answer questions in \\ \verb|day01_activity01_montyhall.ipynb|.

\vspace{\stretch{0.1}}

\homework Following the instructions in \verb|day01_activity02_twoChildren.ipynb|, complete a code for Two-Children puzzle simulation and answer questions


\end{document}
